Different rules for binocular combination of luminance flicker in cortical and subcortical pathways

### 2023-01-10 manuscript.tex
Small edits for Aurelio's comments. Somehow I broke everything after the matching figure caption...

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Different rules for binocular combination of luminance flicker in cortical and subcortical pathways},
  pdfauthor={Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade \& Daniel H. Baker},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Different rules for binocular combination of luminance flicker in cortical and subcortical pathways}
\author{Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade \& Daniel H. Baker}
\date{2023-01-07}

@@

A similar analysis pipeline was adopted for the pupillometry data. The data were converted from mp4 videos to a csv text file using the Pupil Player software (Kassner et al., 2014), which estimated pupil diameter for each eye on each frame using a 3D model of the eyeball. The individual data were then loaded into R for analysis, where again a ten-second waveform for each trial in each eye was extracted (excluding the first two seconds after stimulus onset). We interpolated across any dropped or missing frames to ensure regular and continuous sampling over time. The Fourier transform was calculated for each waveform, and all repetitions of each condition were pooled across eye and then averaged. Finally, data were averaged across all participants to obtain the group results. Again, we used coherent averaging, and excluded outlying data points in the same way as for the EEG data. Note that previous pupillometry studies using luminance flicker have tended to fit a single sine-wave at the fundamental frequency, rather than using Fourier analysis (e.g. Spitschan et al., 2014). The Fourier approach is more robust to noise at other frequencies and has been used in some previous studies (see Barrionuevo et al., 2014; Barrionuevo and Cao, 2016). Additionally, it makes the pupillometry analysis consistent with standard practice in steady-state EEG analysis (e.g. Figueira et al., 2022).

@@

Despite being derived from the model proposed by Meese et al. (2006), the simplifications applied to this architecture make it very similar to other models (e.g. Ding and Sperling, 2006; Doesschate and Alpern, 1967; Legge, 1984; Schrödinger, 1926). In particular we fixed the numerator exponent at 2 in our model, because otherwise this value tends to trade off with the weight of interocular suppression (see Baker et al., 2012; Kingdom and Libenson, 2015). Our key parameter of interest is the weight of interocular suppression. Large values around \(w\) = 1 result in a very small or nonexistent binocular advantage at suprathreshold contrasts, consistent with previous work using grating stimuli (Baker and Wade, 2017). Low values around \(w\) = 0 produce substantial, near-linear binocular facilitation (Baker et al., 2020).

We implemented the model within a Bayesian framework using the Stan software (Carpenter et al., 2017). This allowed us to estimate group-level posterior parameter distributions for the weight of interocular suppression, \(w\), and the other free model parameters \(R_{max}\), \(Z\) and \(n\). The prior distributions for all parameters were Gaussian, with means and standard deviations of 1 and 0.5 for \(w\) and \(R_{max}\), and 5 and 2 for \(Z\) and \(n\). We sampled from a Student's t-distribution for the amplitudes in the pupillometry and EEG experiments, and from a Bernoulli distribution for the single trial matching data. The models were fit using the individual data across all participants, independently for each data set. Because we used coherent averaging across participants, the group average amplitudes are shifted vertically relative to the model predictions, which are based on hierarchical fits to the individual participant amplitudes (put another way, the model does not implement coherent averaging). However it is clear that the character of the model in all cases gives a good representation of the data. We took posterior samples at over a million steps for each data set, using a computer cluster, and retained 10\% of samples for plotting.

@@

When stimuli in both eyes flicker at 2Hz, the binocular responses at the first (Figure \ref{fig:EEGdata}c) and second (Figure \ref{fig:EEGdata}f) harmonics are substantially greater than the monocular responses, particularly at high contrasts. Analysis of variance on the complex values (\(ANOVA^2_{circ}\)) revealed a main effect of contrast (F(8,580) = 4.38, \(p\) \textless{} 0.001) and an interaction effect (F(8,580) = 61.58, \(p\) \textless{} 0.001), but no effect of condition (\(p\) = 0.13). For the cross-frequency conditions (Figure \ref{fig:EEGdata}d,g), there was no appreciable effect of adding a 1.6Hz component on the response at 2Hz or 4Hz (no effect of condition, and no interaction). Similarly, there were no clear interocular interactions between frequencies in the responses at 1.6Hz (Figure \ref{fig:EEGdata}e) and 3.2Hz (Figure \ref{fig:EEGdata}h). This pattern of results suggests that processing of temporal luminance modulations happens in a more linear way in visual cortex (indexed by EEG), compared with subcortical pathways (indexed by pupillometry), and shows no evidence of interocular suppression.

@@

In Experiment 1 we found evidence of stronger binocular facilitation for cortical responses to luminance flicker (measured using EEG), compared with subcortical responses (measured using pupillometry). Since perception is dependent on cortical responses, these results provide a clear prediction for perceived contrast judgements indexed by psychophysical contrast matching paradigms (e.g. Anstis and Ho, 1998; Levelt, 1965; Quaia et al., 2018). We therefore conducted such an experiment, in which participants judged which of two stimuli had the greater perceived amplitude of flicker. One stimulus was a matching stimulus, that had a fixed binocular flicker amplitude of either 24\% or 48\% (temporal) contrast. The other stimulus was a target stimulus, the contrast of which was controlled by a staircase algorithm. We tested 9 ratios of contrast between the left and right eyes.

@@

{\centering \includegraphics[width=0.5\linewidth]{Figures/matchingdata}

\caption{Contrast matching functions. Dotted and dashed lines are predictions of canonical summation models with a linear exponent (dotted) or an infinite exponent (dashed). Error bars indicate the standard error across participants (N=10), and are constrained along radial lines converging at the origin.}\label{fig:matchingdata}
\end{figure}

@@

Pupil size determines the total amount of light falling on the retina. It is therefore the case that fluctuations in pupil diameter have a downstream effect on the signals reaching cortex. We did not incorporate such interactions into our computational model, though in principle this might be worthwhile. However we anticipate than any such effects would be small, since pupil modulations at 2Hz are in the order of 2\% of overall diameter (e.g. Spitschan et al., 2014). It is also the case that cortical activity can modulate pupil diameter, usually through arousal and attention mechanisms (e.g. Bradley et al., 2008). We think it unlikely that these temporally coarse processes would have a differential effect on e.g.~monocular and binocular stimulation conditions in our experiment, and any fluctuations during an experimental session (perhaps owing to fatigue) will be equivalent for our comparisons of interest. Therefore we make the simplifying assumption that the two pathways are effectively distinct, but hope to investigate this directly in future neuroimaging work.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Angee2021}{}}%
Angée C, Nedelec B, Erjavec E, Rozet J-M, Fares Taie L. 2021. Congenital microcoria: Clinical features and molecular genetics. \emph{Genes (Basel)} \textbf{12}. doi:\href{https://doi.org/10.3390/genes12050624}{10.3390/genes12050624}

\leavevmode\vadjust pre{\hypertarget{ref-Anstis1998}{}}%
Anstis S, Ho A. 1998. Nonlinear combination of luminance excursions during flicker, simultaneous contrast, afterimages and binocular fusion. \emph{Vision Res} \textbf{38}:523--39. doi:\href{https://doi.org/10.1016/s0042-6989(97)00167-3}{10.1016/s0042-6989(97)00167-3}

\leavevmode\vadjust pre{\hypertarget{ref-Baker2021}{}}%
Baker DH. 2021. Statistical analysis of periodic data in neuroscience. \emph{Neurons, Behavior, Data analysis, and Theory} \textbf{5}. doi:\href{https://doi.org/10.51628/001c.27680}{10.51628/001c.27680}

\leavevmode\vadjust pre{\hypertarget{ref-Baker2018}{}}%
Baker DH, Lygo FA, Meese TS, Georgeson MA. 2018. Binocular summation revisited: Beyond \(\sqrt{2}\). \emph{Psychol Bull} \textbf{144}:1186--1199. doi:\href{https://doi.org/10.1037/bul0000163}{10.1037/bul0000163}

\leavevmode\vadjust pre{\hypertarget{ref-Baker2007}{}}%
Baker DH, Meese TS, Georgeson MA. 2007. Binocular interaction: Contrast matching and contrast discrimination are predicted by the same model. \emph{Spat Vis} \textbf{20}:397--413. doi:\href{https://doi.org/10.1163/156856807781503622}{10.1163/156856807781503622}

\leavevmode\vadjust pre{\hypertarget{ref-Baker2020}{}}%
Baker DH, Vilidaite G, McClarnon E, Valkova E, Bruno A, Millman RE. 2020. Binaural summation of amplitude modulation involves weak interaural suppression. \emph{Sci Rep} \textbf{10}:3560. doi:\href{https://doi.org/10.1038/s41598-020-60602-5}{10.1038/s41598-020-60602-5}

\leavevmode\vadjust pre{\hypertarget{ref-Baker2017}{}}%
Baker DH, Wade AR. 2017. Evidence for an optimal algorithm underlying signal combination in human visual cortex. \emph{Cereb Cortex} \textbf{27}:254--264. doi:\href{https://doi.org/10.1093/cercor/bhw395}{10.1093/cercor/bhw395}

\leavevmode\vadjust pre{\hypertarget{ref-Baker2012}{}}%
Baker DH, Wallis SA, Georgeson MA, Meese TS. 2012. Nonlinearities in the binocular combination of luminance and contrast. \emph{Vision Res} \textbf{56}:1--9. doi:\href{https://doi.org/10.1016/j.visres.2012.01.008}{10.1016/j.visres.2012.01.008}

\leavevmode\vadjust pre{\hypertarget{ref-Barrionuevo2016}{}}%
Barrionuevo PA, Cao D. 2016. Luminance and chromatic signals interact differently with melanopsin activation to control the pupil light response. \emph{J Vis} \textbf{16(11)}:29. doi:\href{https://doi.org/10.1167/16.11.29}{10.1167/16.11.29}

\leavevmode\vadjust pre{\hypertarget{ref-Barrionuevo2014}{}}%
Barrionuevo PA, Nicandro N, McAnany JJ, Zele AJ, Gamlin P, Cao D. 2014. Assessing rod, cone, and melanopsin contributions to human pupil flicker responses. \emph{Invest Ophthalmol Vis Sci} \textbf{55}:719--27. doi:\href{https://doi.org/10.1167/iovs.13-13252}{10.1167/iovs.13-13252}

\leavevmode\vadjust pre{\hypertarget{ref-Bradley2008}{}}%
Bradley MM, Miccoli L, Escrig MA, Lang PJ. 2008. The pupil as a measure of emotional arousal and autonomic activation. \emph{Psychophysiology} \textbf{45}:602--7. doi:\href{https://doi.org/10.1111/j.1469-8986.2008.00654.x}{10.1111/j.1469-8986.2008.00654.x}

\leavevmode\vadjust pre{\hypertarget{ref-Campbell1965}{}}%
Campbell FW, Green DG. 1965. Monocular versus binocular visual acuity. \emph{Nature} \textbf{208}:191--2. doi:\href{https://doi.org/10.1038/208191a0}{10.1038/208191a0}

\leavevmode\vadjust pre{\hypertarget{ref-Carpenter2017}{}}%
Carpenter B, Gelman A, Hoffman MD, Lee D, Goodrich B, Betancourt M, Brubaker M, Guo J, Li P, Riddell A. 2017. Stan: A probabilistic programming language. \emph{Journal of Statistical Software} \textbf{76}:1--32. doi:\href{https://doi.org/10.18637/jss.v076.i01}{10.18637/jss.v076.i01}

\leavevmode\vadjust pre{\hypertarget{ref-Delorme2004}{}}%
Delorme A, Makeig S. 2004. {EEGLAB}: An open source toolbox for analysis of single-trial {EEG} dynamics including independent component analysis. \emph{J Neurosci Methods} \textbf{134}:9--21. doi:\href{https://doi.org/10.1016/j.jneumeth.2003.10.009}{10.1016/j.jneumeth.2003.10.009}

\leavevmode\vadjust pre{\hypertarget{ref-Ding2006}{}}%
Ding J, Sperling G. 2006. A gain-control theory of binocular combination. \emph{Proc Natl Acad Sci U S A} \textbf{103}:1141--6. doi:\href{https://doi.org/10.1073/pnas.0509629103}{10.1073/pnas.0509629103}

\leavevmode\vadjust pre{\hypertarget{ref-Doesschate1967}{}}%
Doesschate J ten, Alpern M. 1967. Effect of photoexcitation of the two retinas on pupil size. \emph{J Neurophysiol} \textbf{30}:562--76. doi:\href{https://doi.org/10.1152/jn.1967.30.3.562}{10.1152/jn.1967.30.3.562}

\leavevmode\vadjust pre{\hypertarget{ref-Economides2011}{}}%
Economides JR, Sincich LC, Adams DL, Horton JC. 2011. Orientation tuning of cytochrome oxidase patches in macaque primary visual cortex. \emph{Nat Neurosci} \textbf{14}:1574--80. doi:\href{https://doi.org/10.1038/nn.2958}{10.1038/nn.2958}

\leavevmode\vadjust pre{\hypertarget{ref-Edwards1995}{}}%
Edwards DP, Purpura KP, Kaplan E. 1995. Contrast sensitivity and spatial frequency response of primate cortical neurons in and around the cytochrome oxidase blobs. \emph{Vision Res} \textbf{35}:1501--23. doi:\href{https://doi.org/10.1016/0042-6989(94)00253-i}{10.1016/0042-6989(94)00253-i}

\leavevmode\vadjust pre{\hypertarget{ref-Ernst2002}{}}%
Ernst MO, Banks MS. 2002. Humans integrate visual and haptic information in a statistically optimal fashion. \emph{Nature} \textbf{415}:429--33. doi:\href{https://doi.org/10.1038/415429a}{10.1038/415429a}

\leavevmode\vadjust pre{\hypertarget{ref-Figueira2022}{}}%
Figueira JSB, Kutlu E, Scott LS, Keil A. 2022. The FreqTag toolbox: A principled approach to analyzing electrophysiological time series in frequency tagging paradigms. \emph{Dev Cogn Neurosci} \textbf{54}:101066. doi:\href{https://doi.org/10.1016/j.dcn.2022.101066}{10.1016/j.dcn.2022.101066}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Different rules for binocular combination of luminance flicker in cortical and subcortical pathways},
  pdfauthor={Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade \& Daniel H. Baker},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

\title{Different rules for binocular combination of luminance flicker in cortical and subcortical pathways}
\author{Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade \& Daniel H. Baker}
\date{2023-01-10}

@@

A similar analysis pipeline was adopted for the pupillometry data. The data were converted from mp4 videos to a csv text file using the Pupil Player software (Kassner et al., 2014), which estimated pupil diameter for each eye on each frame using a 3D model of the eyeball. The individual data were then loaded into R for analysis, where again a ten-second waveform for each trial in each eye was extracted (excluding the first two seconds after stimulus onset). We interpolated across any dropped or missing frames to ensure regular and continuous sampling over time. The Fourier transform was calculated for each waveform, and all repetitions of each condition were pooled across eye and then averaged. Finally, data were averaged across all participants to obtain the group results. Again, we used coherent averaging, and excluded outlying data points in the same way as for the EEG data. Note that previous pupillometry studies using luminance flicker have tended to fit a single sine-wave at the fundamental frequency, rather than using Fourier analysis (e.g.~Spitschan et al., 2014). The Fourier approach is more robust to noise at other frequencies and has been used in some previous studies (see Barrionuevo et al., 2014; Barrionuevo and Cao, 2016). Additionally, it makes the pupillometry analysis consistent with standard practice in steady-state EEG analysis (e.g.~Figueira et al., 2022).

@@

Despite being derived from the model proposed by Meese et al. (2006), the simplifications applied to this architecture make it very similar to other models (e.g.~Ding and Sperling, 2006; Doesschate and Alpern, 1967; Legge, 1984; Schrödinger, 1926). In particular we fixed the numerator exponent at 2 in our model, because otherwise this value tends to trade off with the weight of interocular suppression (see Baker et al., 2012; Kingdom and Libenson, 2015). Our key parameter of interest is the weight of interocular suppression. Large values around \(w\) = 1 result in a very small or nonexistent binocular advantage at suprathreshold contrasts, consistent with previous work using grating stimuli (Baker and Wade, 2017). Low values around \(w\) = 0 produce substantial, near-linear binocular facilitation (Baker et al., 2020).

We implemented the model within a Bayesian framework using the Stan software (Carpenter et al., 2017). This allowed us to estimate group-level posterior parameter distributions for the weight of interocular suppression, \(w\), and the other free model parameters \(R_{max}\), \(Z\) and \(n\). The prior distributions for all parameters were Gaussian, with means and standard deviations of 1 and 0.5 for \(w\) and \(R_{max}\), and 5 and 2 for \(Z\) and \(n\), with these values chosen based on previous literature (Baker et al., 2012; Meese et al., 2006). We sampled from a Student's t-distribution for the amplitudes in the pupillometry and EEG experiments, and from a Bernoulli distribution for the single trial matching data. The models were fit using the individual data across all participants, independently for each data set. Because we used coherent averaging across participants, the group average amplitudes are shifted vertically relative to the model predictions, which are based on hierarchical fits to the individual participant amplitudes (put another way, the model does not implement coherent averaging). However it is clear that the character of the model in all cases gives a good representation of the data, as it can be seen in figures \ref{fig:modelfigure}a-h. We took posterior samples at over a million steps for each data set, using a computer cluster, and retained 10\% of samples for plotting.

@@

When stimuli in both eyes flicker at 2Hz, the binocular responses at the first (Figure \ref{fig:EEGdata}c) and second (Figure \ref{fig:EEGdata}f) harmonics are substantially greater than the monocular responses, particularly at high contrasts. Analysis of variance on the complex values (\(ANOVA^2_{circ}\)) revealed a main effect of contrast (F(8,580) = 4.38, \(p\) \textless{} 0.001) and an interaction effect (F(8,580) = 61.58, \(p\) \textless{} 0.001), but no effect of condition (\(p\) = 0.13) at the first harmonic, with a similar pattern of statistical results obtained at the second harmonic. For the cross-frequency conditions (Figure \ref{fig:EEGdata}d,g), there was no appreciable effect of adding a 1.6Hz component on the response at 2Hz or 4Hz (no effect of condition, and no interaction). Similarly, there were no clear interocular interactions between frequencies in the responses at 1.6Hz (Figure \ref{fig:EEGdata}e) and 3.2Hz (Figure \ref{fig:EEGdata}h). This pattern of results suggests that processing of temporal luminance modulations happens in a more linear way in visual cortex (indexed by EEG), compared with subcortical pathways (indexed by pupillometry), and shows no evidence of interocular suppression.

@@

In Experiment 1 we found evidence of stronger binocular facilitation for cortical responses to luminance flicker (measured using EEG), compared with subcortical responses (measured using pupillometry). Since perception is dependent on cortical responses, these results provide a clear prediction for perceived contrast judgements indexed by psychophysical contrast matching paradigms (e.g.~Anstis and Ho, 1998; Levelt, 1965; Quaia et al., 2018). We therefore conducted such an experiment, in which participants judged which of two stimuli had the greater perceived amplitude of flicker. One stimulus was a matching stimulus, that had a fixed binocular flicker amplitude of either 24\% or 48\% (temporal) contrast. The other stimulus was a target stimulus, the contrast of which was controlled by a staircase algorithm. We tested 9 ratios of contrast between the left and right eyes.

@@

\textbackslash begin\{figure\}

\{\centering \includegraphics[width=0.5\linewidth]{Figures/matchingdata}

\}

\textbackslash caption\{Contrast matching functions. Dotted and dashed lines are predictions of canonical summation models with a linear exponent (dotted) or an infinite exponent (dashed). Error bars indicate the standard error across participants (N=10), and are constrained along radial lines converging at the origin. Note that, for the 48\% match, the data point on the x axis falls higher than 100\% contrast. This is because some individual results were interpolated above 100\%, therefore skewing the results slighlty above that value. The effects are negligible.\}\label{fig:matchingdata}
\textbackslash end\{figure\}

@@

Pupil size determines the total amount of light falling on the retina. It is therefore the case that fluctuations in pupil diameter have a downstream effect on the signals reaching cortex. We did not incorporate such interactions into our computational model, though in principle this might be worthwhile. However we anticipate than any such effects would be small, since pupil modulations at 2Hz are in the order of 2\% of overall diameter (e.g.~Spitschan et al., 2014). It is also the case that cortical activity can modulate pupil diameter, usually through arousal and attention mechanisms (e.g.~Bradley et al., 2008). We think it unlikely that these temporally coarse processes would have a differential effect on e.g.~monocular and binocular stimulation conditions in our experiment, and any fluctuations during an experimental session (perhaps owing to fatigue) will be equivalent for our comparisons of interest. Therefore we make the simplifying assumption that the two pathways are effectively distinct, but hope to investigate this directly in future neuroimaging work.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Angee2021}{}%
Angée C, Nedelec B, Erjavec E, Rozet J-M, Fares Taie L. 2021. Congenital microcoria: Clinical features and molecular genetics. \emph{Genes (Basel)} \textbf{12}. doi:\href{https://doi.org/10.3390/genes12050624}{10.3390/genes12050624}

\leavevmode\hypertarget{ref-Anstis1998}{}%
Anstis S, Ho A. 1998. Nonlinear combination of luminance excursions during flicker, simultaneous contrast, afterimages and binocular fusion. \emph{Vision Res} \textbf{38}:523--39. doi:\href{https://doi.org/10.1016/s0042-6989(97)00167-3}{10.1016/s0042-6989(97)00167-3}

\leavevmode\hypertarget{ref-Baker2021}{}%
Baker DH. 2021. Statistical analysis of periodic data in neuroscience. \emph{Neurons, Behavior, Data analysis, and Theory} \textbf{5}. doi:\href{https://doi.org/10.51628/001c.27680}{10.51628/001c.27680}

\leavevmode\hypertarget{ref-Baker2018}{}%
Baker DH, Lygo FA, Meese TS, Georgeson MA. 2018. Binocular summation revisited: Beyond \(\sqrt{2}\). \emph{Psychol Bull} \textbf{144}:1186--1199. doi:\href{https://doi.org/10.1037/bul0000163}{10.1037/bul0000163}

\leavevmode\hypertarget{ref-Baker2007}{}%
Baker DH, Meese TS, Georgeson MA. 2007. Binocular interaction: Contrast matching and contrast discrimination are predicted by the same model. \emph{Spat Vis} \textbf{20}:397--413. doi:\href{https://doi.org/10.1163/156856807781503622}{10.1163/156856807781503622}

\leavevmode\hypertarget{ref-Baker2020}{}%
Baker DH, Vilidaite G, McClarnon E, Valkova E, Bruno A, Millman RE. 2020. Binaural summation of amplitude modulation involves weak interaural suppression. \emph{Sci Rep} \textbf{10}:3560. doi:\href{https://doi.org/10.1038/s41598-020-60602-5}{10.1038/s41598-020-60602-5}

\leavevmode\hypertarget{ref-Baker2017}{}%
Baker DH, Wade AR. 2017. Evidence for an optimal algorithm underlying signal combination in human visual cortex. \emph{Cereb Cortex} \textbf{27}:254--264. doi:\href{https://doi.org/10.1093/cercor/bhw395}{10.1093/cercor/bhw395}

\leavevmode\hypertarget{ref-Baker2012}{}%
Baker DH, Wallis SA, Georgeson MA, Meese TS. 2012. Nonlinearities in the binocular combination of luminance and contrast. \emph{Vision Res} \textbf{56}:1--9. doi:\href{https://doi.org/10.1016/j.visres.2012.01.008}{10.1016/j.visres.2012.01.008}

\leavevmode\hypertarget{ref-Barrionuevo2016}{}%
Barrionuevo PA, Cao D. 2016. Luminance and chromatic signals interact differently with melanopsin activation to control the pupil light response. \emph{J Vis} \textbf{16(11)}:29. doi:\href{https://doi.org/10.1167/16.11.29}{10.1167/16.11.29}

\leavevmode\hypertarget{ref-Barrionuevo2014}{}%
Barrionuevo PA, Nicandro N, McAnany JJ, Zele AJ, Gamlin P, Cao D. 2014. Assessing rod, cone, and melanopsin contributions to human pupil flicker responses. \emph{Invest Ophthalmol Vis Sci} \textbf{55}:719--27. doi:\href{https://doi.org/10.1167/iovs.13-13252}{10.1167/iovs.13-13252}

\leavevmode\hypertarget{ref-Bradley2008}{}%
Bradley MM, Miccoli L, Escrig MA, Lang PJ. 2008. The pupil as a measure of emotional arousal and autonomic activation. \emph{Psychophysiology} \textbf{45}:602--7. doi:\href{https://doi.org/10.1111/j.1469-8986.2008.00654.x}{10.1111/j.1469-8986.2008.00654.x}

\leavevmode\hypertarget{ref-Campbell1965}{}%
Campbell FW, Green DG. 1965. Monocular versus binocular visual acuity. \emph{Nature} \textbf{208}:191--2. doi:\href{https://doi.org/10.1038/208191a0}{10.1038/208191a0}

\leavevmode\hypertarget{ref-Carpenter2017}{}%
Carpenter B, Gelman A, Hoffman MD, Lee D, Goodrich B, Betancourt M, Brubaker M, Guo J, Li P, Riddell A. 2017. Stan: A probabilistic programming language. \emph{Journal of Statistical Software} \textbf{76}:1--32. doi:\href{https://doi.org/10.18637/jss.v076.i01}{10.18637/jss.v076.i01}

\leavevmode\hypertarget{ref-Delorme2004}{}%
Delorme A, Makeig S. 2004. EEGLAB: An open source toolbox for analysis of single-trial EEG dynamics including independent component analysis. \emph{J Neurosci Methods} \textbf{134}:9--21. doi:\href{https://doi.org/10.1016/j.jneumeth.2003.10.009}{10.1016/j.jneumeth.2003.10.009}

\leavevmode\hypertarget{ref-Ding2006}{}%
Ding J, Sperling G. 2006. A gain-control theory of binocular combination. \emph{Proc Natl Acad Sci U S A} \textbf{103}:1141--6. doi:\href{https://doi.org/10.1073/pnas.0509629103}{10.1073/pnas.0509629103}

\leavevmode\hypertarget{ref-Doesschate1967}{}%
Doesschate J ten, Alpern M. 1967. Effect of photoexcitation of the two retinas on pupil size. \emph{J Neurophysiol} \textbf{30}:562--76. doi:\href{https://doi.org/10.1152/jn.1967.30.3.562}{10.1152/jn.1967.30.3.562}

\leavevmode\hypertarget{ref-Economides2011}{}%
Economides JR, Sincich LC, Adams DL, Horton JC. 2011. Orientation tuning of cytochrome oxidase patches in macaque primary visual cortex. \emph{Nat Neurosci} \textbf{14}:1574--80. doi:\href{https://doi.org/10.1038/nn.2958}{10.1038/nn.2958}

\leavevmode\hypertarget{ref-Edwards1995}{}%
Edwards DP, Purpura KP, Kaplan E. 1995. Contrast sensitivity and spatial frequency response of primate cortical neurons in and around the cytochrome oxidase blobs. \emph{Vision Res} \textbf{35}:1501--23. doi:\href{https://doi.org/10.1016/0042-6989(94)00253-i}{10.1016/0042-6989(94)00253-i}

\leavevmode\hypertarget{ref-Ernst2002}{}%
Ernst MO, Banks MS. 2002. Humans integrate visual and haptic information in a statistically optimal fashion. \emph{Nature} \textbf{415}:429--33. doi:\href{https://doi.org/10.1038/415429a}{10.1038/415429a}

\leavevmode\hypertarget{ref-Figueira2022}{}%
Figueira JSB, Kutlu E, Scott LS, Keil A. 2022. The freqtag toolbox: A principled approach to analyzing electrophysiological time series in frequency tagging paradigms. \emph{Dev Cogn Neurosci} \textbf{54}:101066. doi:\href{https://doi.org/10.1016/j.dcn.2022.101066}{10.1016/j.dcn.2022.101066}

\leavevmode\hypertarget{ref-Gheiratmand2013}{}%
Gheiratmand M, Meese TS, Mullen KT. 2013. Blobs versus bars: Psychophysical evidence supports two types of orientation response in human color vision. \emph{Journal of Vision} \textbf{13(1)}:2. doi:\href{https://doi.org/10.1167/13.1.2}{10.1167/13.1.2}
